---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(cryptorng)
```

# `{cryptorng}` Generate random bytes from cryptographically secure random number sources

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/cryptorng-dev/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/cryptorng-dev/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{cryptorng}` provides cross-platform [cryptographically secure pseudorandom number generators (CSPRNG)](https://en.wikipedia.org/wiki/Cryptographically_secure_pseudorandom_number_generator).

A good source of random bytes is essential for generating secure encryption 
inputs like keys and nonces (e.g. for [`rmonocypher`](https://github.com/coolbutuseless/rmonocypher)).

The method used for generating random values varies depending on the OS: 

| OS          | CSPRNG                     |
|-------------|----------------------------|
| macOS, *BSD | `arc4random_buf()`         |
| Linux       | `syscall(SYS_getrandom())` |
| Windows     | `BCryptGenRandom()`        |

All these random number generators are internally seeded by the OS using entropy 
gathered from multiple sources and are considered cryptographically secure.


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/cryptorng) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/cryptorng')
```

## Generate random bytes


```{r example}
library(cryptorng)

generate_bytes(16)
generate_bytes(16, type = 'string')
```



# C code

This package and code is MIT licensed.  Please feel free to 
incorporate/adapt this code into your own project.

```
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Get random bytes from the system RNG  (C Callable)
//
// @param buf pre-allocated buffer in which to put the random bytes
// @param n number of bytes.  Note: when a system RNG runs out of entropy
//        it may return fewer bytes than expected. This function throws an 
//        error in this situation
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
void generate_bytes(void *buf, size_t n) {
  
  //~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  // macOS and BSD support arc4random_buf()
  //~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#if defined(__APPLE__) || defined(__FreeBSD__) || defined(__OpenBSD__) || defined(__NetBSD__)
  // void arc4random_buf(void buf[.n], size_t n);
  arc4random_buf(buf, n); 
  
  //~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  // Windows: use BCryptGenRandom
  //~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#elif defined(_WIN32)  
  // NTSTATUS BCryptGenRandom(
  //     [in, out] BCRYPT_ALG_HANDLE hAlgorithm,
  //     [in, out] PUCHAR            pbBuffer,
  //     [in]      ULONG             cbBuffer,
  //     [in]      ULONG             dwFlags
  // );
  // dwFlags = BCRYPT_USE_SYSTEM_PREFERRED_RNG - Use the system-preferred random 
  // number generator algorithm. The hAlgorithm parameter must be NULL. 
  size_t status = (size_t)BCryptGenRandom( NULL, ( PUCHAR ) buf, n, BCRYPT_USE_SYSTEM_PREFERRED_RNG );
  // Return value is 'NTSTATUS' value. STATUS_SUCCESS = 0.
  if (status != 0) {
    error("cryptorng_windows() error: Status = %zu.\n", status);
  }
  
  //~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  // Linux use 'Sys_getrandom()'
  //~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#elif defined(__linux__)
  size_t status = (size_t)syscall( SYS_getrandom, buf, n, 0 );
  if (status < 0 || status != n) {
    error("cryptorng_linux() error: Status = %zu.\n", status);
  }
  
#else
#error no secrure random() implemented for this platform
#endif 
}
```



